A meta analysis of tournaments and an evaluation of performance in the Iterated Prisoner's Dilemma.

### 2019-12-24 paper/main.tex
address minor comments

\begin{abstract}
The Iterated Prisoner's Dilemma has been used for decades as a model of
behavioural interactions. From the celebrated performance of Tit for Tat, to the
introduction of the zero-determinant strategies, to the use of sophisticated
structures such as neural networks, the literature has been exploring the
performance of strategies in the game for years. The results of the literature,
however, have been relying on the performance of specific strategies in a finite
number of tournaments, whereas this manuscript evaluates
\numberofstrategies strategies' effectiveness in more than 40000 tournaments.
The top ranked strategies are presented, and moreover, the impact of features on
their success are analysed using machine learning techniques. The analysis
determines that the cooperating ratio of a
strategy in a given tournament compared to the mean and median cooperator is the
most important feature. The conclusions are distinct for different types of
tournaments. For instance a strategy with a theory of mind would aim to be the mean/median
cooperator in standard tournaments, whereas in tournaments with probabilistic
ending it would aim to cooperate 10\% of the times the median cooperator did.
\end{abstract}

@@

Conceptualising strategies and understanding the best way of playing the game
has been of interest to the scientific community since the formulation of the
game in 1950~\cite{Flood1958}. Following the computer tournaments of Axelrod in the
1980's~\cite{Axelrod1980a, Axelrod1980b}, a strategy's performance in a round
robin computer tournament became a common evaluation technique for newly designed
strategies. Today more than 200 strategies exist in the literature and several
tournaments, excluding Axelrod's, have been undertaken~\cite{Bendor1991,
Harper2017, Kendall2007, Stephens2002, Stewart2012}.

@@

However, further research proved that the strategy had weakness, and more
specifically, it was shown that the strategy suffered in environments with
noise~\cite{Bendor1991, Donninger1986, Molander1985, Hammerstein1984}. This was
mainly due to the strategy's lack of generosity and contrition. The strategy was
quick to punish a defection, and in a noisy environment it could lead to a
repeated cycle of defections and cooperations. Some new strategies, more
robust in tournaments with noise, were soon introduced and became the new
protagonists of the game. These include Nice and Forgiving~\cite{Bendor1991},
Pavlov~\cite{Nowak1993} and Generous Tit For Tat~\cite{Nowak1992}.

In 2004, a $20^{\text{th}}$ Anniversary Iterated Prisoner Dilemma Tournament
took place with 233 entries. This time the winning strategy was not designed on
a reciprocity based approach but on a mechanism of
teams~\cite{J.P.Delahaye1993Lp, J.P.Delahaye1995LIeP, A.Rogers2007Ctpw}. A team
from Southampton University took advantage of the fact that a participant was
allowed to submit multiple strategies. They submitted a total of 60 strategies
that could recognised each other and colluded to increase one members score.
This resulted with three of the strategies to be ranked in the top spots. The
performance of the Southampton University team received mixed attention, though
they had won the tournament as stated in~\cite{us_blog} "technically this
strategy violates the spirit of the Prisoner's Dilemma, which assumes that the
two prisoners cannot communicate with one another".

@@

Though only a select pieces of work have been discussed, the IPD literature is
rich, and new strategies and competitions are being published every year. The
question, however, still remains the same: what is the best way to play the
game? Compared to other works, whereas a few selected strategies are evaluated
on a small number of tournaments, this manuscript evaluates the performance of \numberofstrategies
strategies in \numberofalltournaments tournaments. These tournaments do not
consist by just standard round robin tournaments, but also by tournaments with noise
and tournaments with a probabilistic ending. The later part of the paper, evaluates
the impact of features on the performance of the strategies using modern
machine learning techniques. These features include measures regarding a
strategy's behaviour and measures regarding the tournaments. The data set used
in this work has been made publicly available~\cite{data} and can be used
for further analysis and insights.

The different tournament types as well as the data collection, which is made
possible due an open source package called Axelrod-Python~\cite{axelrodproject},
are covered in Section~\ref{section:data_collection}.
Section~\ref{section:top_performances}, focuses on the best performing
strategies for each type of tournament and overall.
Section~\ref{section:evaluation_of_performance}, explores the traits which
contribute to good performance, and finally the results are summarised in
Section~\ref{section:conclusion}. This manuscripts uses several parameters.
These are introduced in the following sections, however, the full set of
parameters and their definitions are given in Appendix~\ref{app:parameters}.

@@

For the purposes of this manuscript a data set containing results of IPD
tournaments has been generated and is available at~\cite{data}. This was done using the
open source package Axelrod-Python~\cite{axelrodproject}, and more specifically,
version 3.0.0. Axelrod-Python allows for different types of IPD computer
tournaments to be simulated whilst containing a list of over 180 strategies.
Most of these are strategies described in the literature with a few exceptions
being strategies that have been contributed specifically to the package. This
paper make use of \numberofstrategies strategies implemented in version 3.0.0. A
list of the strategies is given in the Appendix~\ref{app:list_of_players}.
Though Axelrod-Python features several tournament types, this work considers
only standard, noisy, probabilistic ending and noisy probabilistic ending
tournaments.

@@

The result summary, Table~\ref{table:output_result}, has \(N\) number of rows
because each row contains information for each strategy that participated in the
tournament. The information includes the strategy's rank, median score, the rate
with which the strategy cooperated $(C_r)$, its match win count and the
probability that the strategy cooperated in the opening move. Moreover, the
probabilities of a strategy being in any of the four states ($CC, CD, DC, DD$),
and the rate of which the strategy cooperated after each state. A measure that
has been manually included is the \textbf{normalised rank}. The normalised rank,
denoted as $r$, is calculated as a strategy's rank divided by the tournament's
size ($N$). In the next section the performance of these strategies is evaluated
based on their normalised rank.

@@

Each strategy participated in multiple tournaments of the same type (on average 5154).
For example Tit For Tat has participated in a total of 5114
tournaments of each type. The strategy's normalised rank distribution in these
is given in Figure~\ref{fig:tit_for_tat_r_distribution}. A value of \(r =
0\) corresponds to a strategy winning the tournament where a value of
\(r = 1\) corresponds to the strategy coming last. Because of the strategies'
multiple entries their performance is evaluated based on the
\textbf{median normalised rank} denoted as \(\bar{r}\).

@@

In standard tournaments 10 out of the 15 top strategies are introduced
in~\cite{Harper2017}. These are strategies based on finite state automata (FSM),
hidden markov models (HMM), artificial neural networks (ANN), lookup tables
(LookerUp) and stochastic lookup tables (Gambler) that have been trained using
reinforcement learning algorithms (evolutionary and particle swarm algorithms).
They have been trained to perform well against the strategies
in~\cite{axelrodproject} in a standard tournament, thus their performance in the
specific setting was anticipated. DoubleCrosser, and Fool Me Once, are
strategies not from the literature but from~\cite{axelrodproject}. DoubleCrosser
is a strategy that makes use of the number of turns because is set to defect on
the last two rounds. The strategy was expected to not perform as well in
tournaments where the number of turns is not specified, but the strategy did not
perform well in tournaments with noise either. Finally, Winner
12~\cite{mathieu2017} and DBS~\cite{Au2006} are both from the the literature.
DBS is strategy specifically designed for noisy environments, however, it ranks
highly only in standard ones.

Figure~\ref{fig:std_results} gives the distributions of $r$ for the top
ranked strategies. The distributions are skewed towards zero and the highest
median, of the top 15 strategies, is at 0.075. This indicates that the top ranked
strategies perform well in any given standard tournament, despite the opponents and
the number of turns.

@@

The top strategies in noisy tournaments are shown in Figure~\ref{fig:noisy_results}. These include deterministic strategies, such
as Tit For 2 Tats~\cite{Axelrod1980b}, Slow Tit For Two Tats~\cite{axelrodproject}, Hard Tit For 2 Tats~\cite{Stewart2012}
and Cycler CCCCCD, and strategies which decide their actions based on the
cooperations to defections ratio, such as ShortMem~\cite{Carvalho2013}, Grumpy
and $e$~\cite{axelrodproject}. Slow Tit For Two Tats is the same strategy as 
Tit For 2 Tats, and at the time of writing this manuscript the
contributors of~\cite{axelrodproject} made a new release where the strategy
has been removed. However, for the purpose of this work the strategy is kept.
The Retaliate and
Limited Retaliate strategies are implemented in~\cite{axelrodproject} by the
same contributor. They are strategies designed to defect if the opponent has
tricked them more often than \(x\%\) of the times that they have
done the same. Finally, in $4^{\text{th}}$ and $9^{\text{th}}$ place are Hunter
strategies which trying to extort, equivalently, strategies that play cyclically
and defectors.

@@

The distributions of \(r\) for the top 6 strategies in probabilistic ending
tournaments over \(p_e\) are given in Figure~\ref{fig:effect_of_probend}.
Figure~\ref{fig:effect_of_probend} shows that the 6 strategies start of with a
high median rank, however, their ranked decreased as the the probability of the
game ending increased and at the point of \(p_e=0.1\) they became the dominant
strategies in their respective tournaments. In essence, what is demonstrated is
that defecting strategies did better when the likelihood of the game ending in
the next turn increased, which is inline with the Folk
Theorem~\cite{Fudenberg2009}. If tournaments where the probability of the game
ending was less than 0.1 were considered then the the top ranked spots are not
dominated by just defecting strategies anymore,
Table~\ref{table:subset_probend_result}. Instead the effective strategies are
now the Meta strategies, trained strategies, Grudger~\cite{axelrodproject} and
Spiteful Tit for Tat~\cite{prison}.

@@

The top ranks include strategies that have been previously mentioned. The
set of Retaliate strategies occupy the top spots followed by BackStabber
and DoubleCrosser. The distributions of the Retaliate strategies have no
statistical difference. Thus, in an IPD tournament where the type is not
specified, playing as any of the Retaliate strategies will have the result.
DoubleCrosser performed well in standard tournaments and the
strategy is just an extension of BackStabber. It should be noted that these
strategies can be characterised as "cheaters". The source code of the strategies
allows them to known the number of turns in a match (if they are specified).
PSO Gambler and Evolved HMM 5 are
trained strategies introduced in~\cite{Harper2017} and Nice Meta Winner and NMWE
Memory One are strategies based on teams. Grudger is a strategy from Axelrod's
original tournament and Forgetful Fool Me Once is based on the same approach as
Grudger. Overall the top 15 strategies are fundamentally different. Some are cheaters,
some are complex, others are simple deterministic strategies and strategies based
on teams. The results of \numberofalltournaments tournaments used in this work imply the following:
they is not a single type of strategy which can performance well in any IPD interaction.

@@

The results of this section imply that successful strategies for specific
settings exist for an IPD tournament. The top ranked strategies in both standard
tournaments and tournaments with probabilistic ending, managed to rank in the
top 10\% of the tournament most of the times.
Strategies in noisy
environments demonstrated that no strategy can be consistently
successful, expected if the value of noise is constrained to less than a half.
Overall, there has been not a single strategy that has shown to perform well in
more than one setting. The aim of the next section is to understand which are
the factors that made these strategies successful, in each setting separately
but also overall.

@@

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[h]
    \begin{center}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{gcgcgc}
    \toprule
    measure & measure explanation &  source & value type & min value & max value \\
    \midrule
stochastic  &  If a strategy is stochastic & strategy classifier from~\cite{axelrodproject} & boolean  & Na &  Na \\
makes use of game &  If a strategy makes used of the game information & strategy classifier from~\cite{axelrodproject} & boolean  & Na &  Na \\
makes use of length &  If a strategy makes used of the number of turns & strategy classifier from~\cite{axelrodproject} & boolean  & Na &  Na \\
memory usage &  The memory size of a strategy divided by the number of turns & memory size from~\cite{axelrodproject} & float & 0 &  1 \\
SSE & A measure of how far a strategy is from ZD behaviour & method described in~\cite{Knight2019} & float & 0 & 1 \\
max cooperating rate $(C_{\text{max}})$  & The biggest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
min cooperating rate $(C_{\text{min}})$ & The smallest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
median cooperating rate $(C_{\text{median}})$ & The median cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
mean cooperating rate $(C_{\text{mean}})$ & The mean cooperating rate in a given tournament  & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{max}}$ & A strategy's cooperating rate divided by the maximum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{min}}$ & A strategy's cooperating rate divided by the minimum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{median}}$ & A strategy's cooperating rate divided by the median  & result summary  & float & 0 & 1\\
$C_r$ / $C_{\text{mean}}$ & A strategy's cooperating rate divided by the mean & result summary  & float & 0 & 1 \\
$C_r$ & The cooperating ratio of a strategy & result summary  & float & 0 & 1 \\
$CC$ to $C$ rate & The probability a strategy will cooperate after a mutual cooperation & result summary  & float & 0 & 1\\
$CD$ to $C$ rate & The probability a strategy will cooperate after being betrayed by the opponent & result summary  & float & 0 & 1 \\
$DC$ to $C$ rate & The probability a strategy will cooperate after betraying the opponent & result summary  & float & 0 & 1 \\
$DD$ to $C$ rate & The probability a strategy will cooperate after a mutual defection & result summary  & float & 0 & 1 \\
$p_n$ & The probability of a player's action being flip at each interaction & trial summary & float & 0 & 1 \\
$n$ & The number of turns & trial summary & integer & 1 & 200 \\
$p_e$ & The probability of a match ending in the next turn & trial summary & float & 0 & 1 \\
$N$ & The number of strategies in the tournament & trial summary & integer & 3 & 195 \\
$k$ & The number of repetitions of a given tournament & trial summary & integer & 10 & 100 \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{The measures which are included in the performance evaluation analysis.}
    \label{table:manual_measures}
\end{table}

@@

The performances are clustered based on the normalised rank. More specifically,
they are clustered 3 times into 2 different clusters based on  on whether their
normalised rank was in the top 5\%, 25\% or 50\% respectively. A random forest
approach~\cite{breiman2001} is then applied to each performance to predict the cluster to
which it has been assigned to. The random forest method
constructs many individual decision trees and the predictions from all trees are
pooled to make the final prediction. The random forest models are trained on a
training set of 70\% of the tournaments results. The accuracy of each model
based on $R^2$ are given by Table~\ref{table:accuracy_random_forest}. The out of
the bag error~\cite{hastie2005} has also been calculated. The models fit well,
and a high value of both the accuracy measure on the test data and the OOB error
indicate that the model is not over fitting.

@@

\begin{figure}[!htbp]
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/standard/_feature_importance_bar_plot_cluster_on_0_05.pdf}
        \end{center}
        \caption{Importance of features for clusters on 5\% performance.}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/standard/_feature_importance_bar_plot_cluster_on_0_25.pdf}
        \end{center}
        \caption{Importance of features for clusters on 25\% performance.}
    \end{subfigure}
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/standard/_feature_importance_bar_plot_cluster_on_0_5.pdf}
        \end{center}
        \caption{Importance of features for clusters on 50\% performance.}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../output/standard/_feature_importance_bar_plot.pdf}
        \end{center}
        \caption{Importance of features for clusters based on \(k\)means algorithm.}
    \end{subfigure}
    \caption{Importance of features in standard tournaments for different
    clustering methods.}\label{fig:clustering_importance_standard}
\end{figure}

\begin{figure}[!htbp]
    \begin{subfigure}{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/noise/_feature_importance_bar_plot_cluster_on_0_05.pdf}
        \end{center}
        \caption{Importance of features for clusters on 5\% performance.}
    \end{subfigure}\hfill
    \begin{subfigure}{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/noise/_feature_importance_bar_plot_cluster_on_0_25.pdf}
        \end{center}
        \caption{Importance of features for clusters on 25\% performance.}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/noise/_feature_importance_bar_plot_cluster_on_0_5.pdf}
        \end{center}
        \caption{Importance of features for clusters on 50\% performance.}
    \end{subfigure}\hfill
    \begin{subfigure}{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../output/noise/_feature_importance_bar_plot.pdf}
        \end{center}
        \caption{Importance of features for clusters based on \(k\)means algorithm.}
    \end{subfigure}
    \caption{Importance of features in noisy tournaments for different
    clustering methods.}\label{fig:clustering_importance_noise}
\end{figure}

>>>>

\begin{abstract}
The Iterated Prisoner's Dilemma has been used for decades as a model of
behavioural interactions. From the celebrated performance of Tit for Tat, to the
introduction of the zero-determinant strategies, to the use of sophisticated
structures such as neural networks, the literature has been exploring the
performance of strategies in the game for years. The results of the literature,
however, have been relying on the performance of specific strategies in a finite
number of tournaments. This manuscript evaluates
\numberofstrategies strategies' effectiveness in more than 40000 tournaments.
The top ranked strategies are presented, and moreover, the impact of features on
their success are analysed using machine learning techniques. The analysis
determines that the cooperation ratio of a
strategy in a given tournament compared to the mean and median cooperator is the
most important feature. The conclusions are distinct for different types of
tournaments. For instance a strategy with a theory of mind would aim to be the mean/median
cooperator in standard tournaments, whereas in tournaments with probabilistic
ending it would aim to cooperate 10\% of the times the median cooperator did.
\end{abstract}

@@

Conceptualising strategies and understanding the best way of playing the game
has been of interest to the scientific community since the formulation of the
game in 1950~\cite{Flood1958}. Following the computer tournaments of Axelrod in the
1980's~\cite{Axelrod1980a, Axelrod1980b}, a strategy's performance in a round
robin computer tournament became a common evaluation technique for newly designed
strategies. Today more than 200 strategies exist in the literature and several
tournaments, following on from Axelrod's, have been undertaken~\cite{Bendor1991,
Harper2017, Kendall2007, Stephens2002, Stewart2012}.

@@

However, further research proved that the strategy had weaknesses, and more
specifically, it was shown that the strategy suffered in environments with
noise~\cite{Bendor1991, Donninger1986, Molander1985, Hammerstein1984}. This was
mainly due to the strategy's lack of generosity and contrition. The strategy was
quick to punish a defection, and in a noisy environment it could lead to a
repeated cycle of defections and cooperations. Some new strategies, more
robust in tournaments with noise, were soon introduced and became the new
protagonists of the game. These include Nice and Forgiving~\cite{Bendor1991},
Pavlov~\cite{Nowak1993} and Generous Tit For Tat~\cite{Nowak1992}.

In 2004, a $20^{\text{th}}$ Anniversary Iterated Prisoner Dilemma Tournament
took place with 233 entries. This time the winning strategy was not designed on
a reciprocity based approach but on a mechanism of
teams~\cite{J.P.Delahaye1993Lp, J.P.Delahaye1995LIeP, A.Rogers2007Ctpw}. A team
from Southampton University took advantage of the fact that a participant was
allowed to submit multiple strategies. They submitted a total of 60 strategies
that could recognise each other and colluded to increase one members score.
This resulted with three of the strategies to be ranked in the top spots. The
performance of the Southampton University team received mixed attention, though
they had won the tournament as stated in~\cite{us_blog} "technically this
strategy violates the spirit of the Prisoner's Dilemma, which assumes that the
two prisoners cannot communicate with one another".

@@

Though only a select pieces of work have been discussed, the IPD literature is
rich, and new strategies and competitions are being published every year. The
question, however, still remains the same: what is the best way to play the
game? Compared to other works, whereas a few selected strategies are evaluated
on a small number of tournaments, this manuscript evaluates the performance of \numberofstrategies
strategies in \numberofalltournaments tournaments. These tournaments do not
consist of just standard round robin tournaments, but also tournaments with noise
and tournaments with a probabilistic ending. The later part of the paper, evaluates
the impact of features on the performance of the strategies using modern
machine learning techniques. These features include measures regarding a
strategy's behaviour and measures regarding the tournaments. The data set used
in this work has been made publicly available~\cite{data} and can be used
for further analysis and insights.

The different tournament types as well as the data collection, which is made
possible due to an open source package called Axelrod-Python~\cite{axelrodproject},
are covered in Section~\ref{section:data_collection}.
Section~\ref{section:top_performances}, focuses on the best performing
strategies for each type of tournament and overall.
Section~\ref{section:evaluation_of_performance}, explores the traits which
contribute to good performance, and finally the results are summarised in
Section~\ref{section:conclusion}. This manuscripts uses several parameters.
These are introduced in the following sections, however, the full set of
parameters and their definitions are given in Appendix~\ref{app:parameters}.
A list of all strategies including citations of their origins is given in
Appendix~\ref{app:list_of_players}.

@@

For the purposes of this manuscript a data set containing results of IPD
tournaments has been generated and is available at~\cite{data}. This was done using the
open source package Axelrod-Python~\cite{axelrodproject}, and more specifically,
version 3.0.0. Axelrod-Python allows for different types of IPD computer
tournaments to be simulated whilst containing a list of over 180 strategies.
Most of these are strategies described in the literature with a few exceptions
being strategies that have been contributed specifically to the package. This
paper make use of \numberofstrategies strategies implemented in version 3.0.0. A
list of the strategies is given in the Appendix~\ref{app:list_of_players}.
Though Axelrod-Python features several tournament types, this work considers
standard, noisy, probabilistic ending and noisy probabilistic ending
tournaments.

@@

The result summary, Table~\ref{table:output_result}, has \(N\) rows
because each row contains information for each strategy that participated in the
tournament. The information includes the strategy's rank, median score, the rate
with which the strategy cooperated $(C_r)$, its match win count and the
probability that the strategy cooperated in the opening move. Moreover, the
probabilities of a strategy being in any of the four states ($CC, CD, DC, DD$),
and the rate of which the strategy cooperated after each state. A measure that
has been manually included is the \textbf{normalised rank}. The normalised rank,
denoted as $r$, is calculated as a strategy's rank divided by the tournament's
size ($N$). In the next section the performance of these strategies is evaluated
based on their normalised rank.

@@

Each strategy participated in multiple tournaments of the same type (on average 5154).
For example Tit For Tat participated in a total of 5114
tournaments of each type. The strategy's normalised rank distribution in these
is given in Figure~\ref{fig:tit_for_tat_r_distribution}. A value of \(r =
0\) corresponds to a strategy winning the tournament where a value of
\(r = 1\) corresponds to the strategy coming last. Because of the strategies'
multiple entries their performance is evaluated based on the
\textbf{median normalised rank} denoted as \(\bar{r}\).

@@

In standard tournaments 10 out of the 15 top strategies are introduced
in~\cite{Harper2017}. These are strategies based on finite state automata (FSM),
hidden markov models (HMM), artificial neural networks (ANN), lookup tables
(LookerUp) and stochastic lookup tables (Gambler) that have been trained using
reinforcement learning algorithms (evolutionary and particle swarm algorithms).
They have been trained to perform well against the strategies
in~\cite{axelrodproject} in a standard tournament, thus their performance in the
specific setting was anticipated. DoubleCrosser, and Fool Me Once, are
strategies not from the literature but from~\cite{axelrodproject}. DoubleCrosser
is a strategy that makes use of the number of turns because it is set to defect on
the last two rounds. The strategy was expected to not perform as well in
tournaments where the number of turns is not specified, but the strategy did not
perform well in tournaments with noise either. Finally, Winner
12~\cite{mathieu2017} and DBS~\cite{Au2006} are both from the the literature.
DBS is strategy specifically designed for noisy environments, however, it ranks
highly only in standard ones.

Figure~\ref{fig:std_results} gives the distributions of $r$ for the top
ranked strategies. The distributions are skewed towards zero and the highest
median, of the top 15 strategies, is at 0.075. This indicates that the top ranked
strategies perform well in any given standard tournament, regardless of the opponents and
the number of turns.

@@

The top strategies in noisy tournaments are shown in Figure~\ref{fig:noisy_results}. These include deterministic strategies, such
as Tit For 2 Tats~\cite{Axelrod1980b}, Slow Tit For Two Tats~\cite{axelrodproject}, Hard Tit For 2 Tats~\cite{Stewart2012}
and Cycler CCCCCD, and strategies which decide their actions based on the
cooperations to defections ratio, such as ShortMem~\cite{Carvalho2013}, Grumpy
and $e$~\cite{axelrodproject}. Slow Tit For Two Tats is the same strategy as 
Tit For 2 Tats, and at the time of writing this manuscript the
contributors of~\cite{axelrodproject} made a new release where the strategy
has been removed. However, for the purpose of this work the strategy is kept.
The Retaliate and
Limited Retaliate strategies are implemented in~\cite{axelrodproject} by the
same contributor. They are strategies designed to defect if the opponent has
tricked them more often than \(x\%\) of the times that they have
done the same. Finally, in $4^{\text{th}}$ and $9^{\text{th}}$ place are Hunter
strategies which try to extort strategies that play cyclically
and defectors.

@@

The distributions of \(r\) for the top 6 strategies in probabilistic ending
tournaments over \(p_e\) are given in Figure~\ref{fig:effect_of_probend}.
Figure~\ref{fig:effect_of_probend} shows that the 6 strategies start of with a
high median rank, however, their ranked decreased as the the probability of the
game ending increased and at the point of \(p_e=0.1\) they became the dominant
strategies in their respective tournaments. In essence, what is demonstrated is
that defecting strategies did better when the likelihood of the game ending in
the next turn increased, which is inline with the Folk
Theorem~\cite{Fudenberg2009}. If tournaments where the probability of the game
ending was less than 0.1 were considered then the the top ranked spots are not
dominated by just defecting strategies anymore, as shown in
Table~\ref{table:subset_probend_result}. Instead the effective strategies are
now the Meta strategies, trained strategies, Grudger~\cite{axelrodproject} and
Spiteful Tit for Tat~\cite{prison}.

@@

The top ranks include strategies that have been previously mentioned. The
set of Retaliate strategies occupy the top spots followed by BackStabber
and DoubleCrosser. The distributions of the Retaliate strategies have no
statistical difference. Thus, in an IPD tournament where the type is not
specified, playing as any of the Retaliate strategies seems like a good approach.
DoubleCrosser performed well in standard tournaments and the
strategy is just an extension of BackStabber. It should be noted that these
strategies can be characterised as ``cheaters". The source code of the strategies
allows them to known the number of turns in a match (if they are specified).
PSO Gambler and Evolved HMM 5 are
trained strategies introduced in~\cite{Harper2017} and Nice Meta Winner and NMWE
Memory One are strategies based on teams. Grudger is a strategy from Axelrod's
original tournament and Forgetful Fool Me Once is based on the same approach as
Grudger. Overall the top 15 strategies are fundamentally different. Some are cheaters,
some are complex, others are simple deterministic strategies and strategies based
on teams. The results of \numberofalltournaments tournaments used in this work imply the following:
there is not a single type of strategy which can performance well in any IPD interaction.

@@

The results of this section imply that successful strategies for specific
settings exist for an IPD tournament. The top ranked strategies in both standard
tournaments and tournaments with probabilistic ending, managed to rank in the
top 10\% of the tournament most of the times.
Strategies in noisy
environments demonstrated that no strategy can be consistently
successful, except if the value of noise is constrained to less than a half.
Overall, there has been not a single strategy that has shown to perform well in
more than one setting. The aim of the next section is to understand the features
that made these strategies successful, in each setting separately
but also overall.

@@

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[h]
    \begin{center}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{gcgcgc}
    \toprule
    measure & measure explanation &  source & value type & min value & max value \\
    \midrule
stochastic  &  If a strategy is stochastic & strategy classifier from~\cite{axelrodproject} & boolean  & Na &  Na \\
makes use of game &  If a strategy makes used of the game information & strategy classifier from~\cite{axelrodproject} & boolean  & Na &  Na \\
makes use of length &  If a strategy makes used of the number of turns & strategy classifier from~\cite{axelrodproject} & boolean  & Na &  Na \\
memory usage &  The memory size of a strategy divided by the number of turns & memory size from~\cite{axelrodproject} & float & 0 &  1 \\
SSE & A measure of how far a strategy is from ZD behaviour & method described in~\cite{Knight2019} & float & 0 & 1 \\
max cooperating rate $(C_{\text{max}})$  & The biggest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
min cooperating rate $(C_{\text{min}})$ & The smallest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
median cooperating rate $(C_{\text{median}})$ & The median cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
mean cooperating rate $(C_{\text{mean}})$ & The mean cooperating rate in a given tournament  & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{max}}$ & A strategy's cooperating rate divided by the maximum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{min}}$ & A strategy's cooperating rate divided by the minimum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{median}}$ & A strategy's cooperating rate divided by the median  & result summary  & float & 0 & 1\\
$C_r$ / $C_{\text{mean}}$ & A strategy's cooperating rate divided by the mean & result summary  & float & 0 & 1 \\
$C_r$ & The cooperating ratio of a strategy & result summary  & float & 0 & 1 \\
$CC$ to $C$ rate & The probability a strategy will cooperate after a mutual cooperation & result summary  & float & 0 & 1\\
$CD$ to $C$ rate & The probability a strategy will cooperate after being betrayed by the opponent & result summary  & float & 0 & 1 \\
$DC$ to $C$ rate & The probability a strategy will cooperate after betraying the opponent & result summary  & float & 0 & 1 \\
$DD$ to $C$ rate & The probability a strategy will cooperate after a mutual defection & result summary  & float & 0 & 1 \\
$p_n$ & The probability of a player's action being flip at each interaction & trial summary & float & 0 & 1 \\
$n$ & The number of turns & trial summary & integer & 1 & 200 \\
$p_e$ & The probability of a match ending in the next turn & trial summary & float & 0 & 1 \\
$N$ & The number of strategies in the tournament & trial summary & integer & 3 & 195 \\
$k$ & The number of repetitions of a given tournament & trial summary & integer & 10 & 100 \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{The features which are included in the performance evaluation analysis.}
    \label{table:manual_measures}
\end{table}

@@

The performances are clustered based on the normalised rank. More specifically,
they are clustered 3 times into 2 different clusters based on  on whether their
normalised rank was in the top 5\%, 25\% or 50\% respectively. A random forest
approach~\cite{breiman2001} is then applied to each performance to predict the cluster to
which it has been assigned to. The random forest method
constructs many individual decision trees and the predictions from all trees are
pooled to make the final prediction. The random forest models are trained on a
training set of 70\% of the tournaments results. The accuracy of each model
based on $R^2$ are given by Table~\ref{table:accuracy_random_forest}. The out of
the bag error (OOB)~\cite{hastie2005} has also been calculated. The models fit well,
and a high value of both the accuracy measure on the test data and the OOB error
indicate that the model is not over fitting.

@@

\begin{figure}[!htbp]
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/standard/_feature_importance_bar_plot_cluster_on_0_05.pdf}
        \end{center}
        \caption{Importance of features for clusters on 5\% performance.}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/standard/_feature_importance_bar_plot_cluster_on_0_25.pdf}
        \end{center}
        \caption{Importance of features for clusters on 25\% performance.}
    \end{subfigure}
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/standard/_feature_importance_bar_plot_cluster_on_0_5.pdf}
        \end{center}
        \caption{Importance of features for clusters on 50\% performance.}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../k_means_output/standard/_feature_importance_bar_plot.pdf}
        \end{center}
        \caption{Importance of features for clusters based on \(k\)means algorithm.}
    \end{subfigure}
    \caption{Importance of features in standard tournaments for different
    clustering methods.}\label{fig:clustering_importance_standard}
\end{figure}

\begin{figure}[!htbp]
    \begin{subfigure}{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/noise/_feature_importance_bar_plot_cluster_on_0_05.pdf}
        \end{center}
        \caption{Importance of features for clusters on 5\% performance.}
    \end{subfigure}\hfill
    \begin{subfigure}{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/noise/_feature_importance_bar_plot_cluster_on_0_25.pdf}
        \end{center}
        \caption{Importance of features for clusters on 25\% performance.}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../new_output/noise/_feature_importance_bar_plot_cluster_on_0_5.pdf}
        \end{center}
        \caption{Importance of features for clusters on 50\% performance.}
    \end{subfigure}\hfill
    \begin{subfigure}{0.5\textwidth}
        \begin{center}
            \includegraphics[width=.75\linewidth]{../k_means_output/noise/_feature_importance_bar_plot.pdf}
        \end{center}
        \caption{Importance of features for clusters based on \(k\)means algorithm.}
    \end{subfigure}
    \caption{Importance of features in noisy tournaments for different
    clustering methods.}\label{fig:clustering_importance_noise}
\end{figure}
